Activities

### 2025-09-18 activities.qmd
after Roes 2025

@@

* Invited session at the [ROeS 2025](https://roes2025.at/index.html) (Conference of the Austro-Swiss Region of the International Biometric Society) in Graz, A, 14-18 Sep 2025. Topic of the session: Contrasting Bayesian and frequentist model building for descriptive research questions - a paired-design experiment of STRATOS-TG2. Invited speaker: Mariana Nold, further speakers for TG2: Michael Kammer, Georg Heinze, Aliaksandr Hubin

>>>>

* Invited session at the [ROeS 2025](https://roes2025.at/index.html) (Conference of the Austro-Swiss Region of the International Biometric Society) in Graz, A, 14-18 Sep 2025. Topic of the session: Contrasting Bayesian and frequentist model building for descriptive research questions - a paired-design experiment of STRATOS-TG2. Invited speaker: [Mariana Nold](/assets/Mariana_slides_Roes_2025.pdf). Further speakers for TG2: [Michael Kammer](/assets/20250902_Kammer.pdf), [Aliaksandr Hubin](/assets/Aliaksandr_slides_Roes_2025.pdf), [Georg Heinze](/assets/2025-09 Roes Heinze.pdf).

<details>
  <summary>Session abstract and slides</summary>

### Abstract

Statistical inference is characterized by a rich epistemological diversity, underpinned by both historical and contemporary debates. The frequentist approach, never having been a closed theory, was characterized by conflicts within statistical methodologies. Similarly, significant differences and even conflicts have existed both within and between frequentist and Bayesian schools.
Despite these tensions, efforts have been made to combine the strengths of frequentist and Bayesian approaches in a pragmatic manner, rather than adhering dogmatically to one paradigm. Our session aims to explore the extent to which modelling approaches are inherently frequentist or Bayesian, and whether these distinctions lead to divergent results. We will compare Bayesian and frequentist model building in four case studies: predicting bacteremia, estimating the proportion of body fat, describing the orbits of exoplanets, investigating the sense of school-belonging in Austrian pupils. For each case study, a research question suited to a multivariable regression model was defined. Two analysis plans were developed per study by different individuals: one Bayesian and one frequentist. Data is currently being analyzed in parallel, and results will be presented in the session. We will thoroughly discuss each step, examining the nuances of each approach to enhance understanding of the methodological and practical implications of statistical inference paradigms in complex, real-world problems.

Pictures:

[Starting the session](/assets/1758101359311.jpeg) (thanks to Sereina Herzog)

[After the session](/assets/PXL_20250918_080447284.jpg) (thanks to Alexandra Kaider)

</details>

@@

* summer activities completed! we are preparing for 2026...
